## Homework 3

<!--
You can use the example3.Rmd or other files to find examples
of the codes you'll need to use to complete this assignment. I strongly
urge you to copy and paste relevant code that you know works from my
examples or instructions and then edit that code so you're working
with the appropriate variables for the homework.
-->

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->

<!--
Read in the homework3.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
-->
The first step is to read in the data.
```{r}
dat<-read.csv("homework3.csv")
```

<!--
Check the structure of the data.
-->
The next step is to check out the structure

```{r}
str(dat)
```
As we can see dat is a data frame with three variables: major,
year, and conservatism.

<!--
Find the mean, median, and standard deviation using the sd() function
for the overall level of conservatism.
-->
Now we'll find some basic summary statistics including the
mean, median, and standard deviation using the sd() function for the overall level of conservatism.

```{r}
mean(dat$conservatism)
median(dat$conservatism)
sd(dat$conservatism)
```
From the previous analyses, we can see that the mean level of
conservation is 52.29458, the median is  53.83231, and the standard deviation is  19.75757.
<!--
Find the mean level of conservation for freshmen only.
-->
Now, let's see what the mean level of conservatism is for freshmen.

```{r}
mean(dat$conservatism[dat$year=="freshman"])
```
Here we can see that the mean for freshmen is 47.50033.

<!--
Find the mean level of conservatism for sophomore business students only.
-->

```{r}
mean(dat$conservatism[dat$year=="sophomore" & dat$major=="business"])
```
Here we can see that the mean for sophmore business students is 61.60146.