## Homework 6

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->

```{r}
library(ggplot2)
library(gplots)
library(dplyr)
```

<!--
Read in the homework6.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
-->

The fabricated data are suppose to approximate how people perform in different types of visual searches. The design of this data is to compare performances in feature searches and conjunction searches. 

```{r}
dat<-read.csv("Homework6.csv")
```

<!--
Check the structure of the data and change any variables that should
be factor variables into factor variables
-->

The next step is to check out the structure

```{r}
str(dat)
```
<!--
Perform a one-way between-subjects ANOVA examining for effects of 
the number of distractors for the comjunction searches only.
Also create a line graph to help you interpret the results of your
analysis.
-->

From the structure we can see that distractors, subject, and subject2
should all be factor variables. So let's fix that.
First we need to get only the data from the conjunction searches.

```{r}
dat$distractor<-as.factor(dat$distractors)
dat$subject<-as.factor(dat$subject)
dat$subject2<-as.factor(dat$subject2)

temp<-dat[dat$type=="conjunction",]%>%group_by(distractors)%>%
  summarize(means=mean(rt),
            sems=sd(rt)/sqrt(length(rt)))

f<-ggplot(temp, aes(x=as.factor(distractors),
                    y=means,
                    group=1))+
  geom_line(size=1)+
  geom_point(size=2)+
  geom_errorbar(aes(ymax=means+sems,
                    ymin=means-sems),
                width=.2)+
  ggtitle("RT for Searches as a Function of Distractors")+
  labs(x="Distractors",y="RT")+
  theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
  theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
  theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
  theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
  theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())
f
```

```{r}
summary(aov(rt~distractors,data=dat[dat$type=="conjunction",]))
```

<!--
Perform a one-way within subjects ANOVA on the same data you analyzed
earlier. Tip: use subject as the subjects variable for this analysis.
-->

```{r}
summary(aov(rt~distractors+Error(subject/distractors),data=dat[dat$type=="conjunction",]))
```

<!--
Perform a between-subjects ANOVA examining for the effects of and
interactions between list position and whether or not the
participants had an intervening distractor task before recalling
the lists of words.
-->

Now, I'll perform a between-subjects 2-way ANOVA looking for effects
of and interactions between distractors and type.

```{r}
 summary(aov(rt~distractors*type,data=dat))
```

<!--
Create a line graph and a grouped bar graph to help us interpret
the results of the previous analyses. Include error bars in your
figures.
-->

To help me interpret the earlier analyses, I'll create line and
grouped bar graphs for the data I analyzed. But first I need to
summarize the data.

```{r}
temp<-dat%>%group_by(distractors,type)%>%
  summarize(means=mean(rt),
            sems=sd(rt)/sqrt(length(rt)))


f<-ggplot(temp,aes(x=distractors,
                    y=means,
                    group=type,
                    color=type))+
  geom_line(size=1)+
  geom_point(size=2)+
  scale_color_manual(values=c("darkgreen","red"),
                     name="Type",
                     breaks=c("feature","conjunction"),
                     labels=c("Feature", "Conjunction"))+
  geom_errorbar(aes(ymax=means+sems,ymin=means-sems),width=.2)+
  ggtitle("RT for Searches as a Function of Distractors and Type")+
  labs(x="Distructors",y="RT")+
  theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
  theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
  theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
  theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
  theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())+
  theme(legend.position=c(.4,.76))+
  theme(legend.background=element_blank())+
  theme(legend.background=element_rect(color="black"))+
  theme(legend.title=element_blank())+
  theme(legend.title=element_text(size=12))+
  theme(legend.title.align=.5)+
  theme(legend.text=element_text(size=10,face="bold"))
f

f<-ggplot(temp,aes(x=distractors,
                    y=means,
                    fill=type))+
  geom_bar(stat="identity",position=position_dodge())+
  scale_fill_manual(values=c("darkgreen","red"),
                     name="Type of Search",
                     breaks=c("feature","conjunction"),
                     labels=c("Feature", "Conjunction"))+
  geom_errorbar(aes(ymax=means+sems,
                    ymin=means-sems),
                width=.2,
                position=position_dodge(.9))+
  ggtitle("RT for Searches as a Function of Distractors and Types")+
  labs(x="Distractors",y="RT")+
  theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
  theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
  theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
  theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
  theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())+
  theme(legend.position=c(.4,.76))+
  theme(legend.background=element_blank())+
  theme(legend.background=element_rect(color="black"))+
  theme(legend.title=element_blank())+
  theme(legend.title=element_text(size=12))+
  theme(legend.title.align=.5)+
  theme(legend.text=element_text(size=10,face="bold"))
f
```

<!--
Interpret the statistical results in light of the figures.
-->

From the figures and earlier analyses we can see that people generally recognize the object they are looking for, for example a red "O" in a group of green "O"s or red "X"s, better when the distractor is just a conjunction instead of a feature. We can also see that in general participants performed worse when they had the feature distractor. Most importantly, the interaction revealed that participants had a quicker response time when exposed to the conjunction. The response times for conjunctions had a faster RT than witht the feature.

<!--
Perform a within-subjects two-way ANOVA examining for the effects
of and interactions between the number of distracts and search type. Tip:
use subject as the subject's variable for this analysis. Interpret
the results of the analysis based on the statistical results and
your earlier figures.
-->

Here I'm performing a within-subjects two-way ANOVA to examine for
the effects of and interactions between distractors and type.

```{r}
summary(aov(rt~distractors*type+
              Error(subject/(distractors*type)),data=dat))
```

From the analyses and the figures we can see that generally people
have a shorter response time with the conjunction rather than the feature, F(1,7) =
1626, p < .05. Finally, we can see that those without a feature perform best for the
words at the beginning and end of the lists while, those with the conjunction perform best for the response time, F(1,7) = 1624, p < .05.

<!--
Finally conduct and interpret the results of a mixed ANOVA in which
number of distractors varies within subjects and type of search varies between
subjects. Make sure to interpret your results. Tip: use subject2
as your subject's variable for this analysis.
-->

Finally, I'll conduct a mixed ANOVA.

```{r}
summary(aov(rt~distractors*type+
              Error(subject2/distractors),
            data=dat))
```

From our anlyses, we can see that individuals without features performed better than those who had, F(1,14) = 182.4, p = < 0.5, which is a significant result.
Additionally, participants generally performed better with the conjuction variable, F(1,14) and (182.4) = 182.4 than with the feature variable (1.14) F=118 with p < .05.
